Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-08-17 manuscript/revision2/revision2.tex
revison2: finalize manuscript and cover letter for resubmission

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, rank-changing
genotype-by-environment interaction, photoperiod, cumulative
rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author[a%
,\equalcont%
,\correspond%
]{Alice H. MacQueen}
\author[a%
,\equalcont%
%
]{Li Zhang}
\author[a%
,\equalcont%
%
]{Samuel A. Smith}
\author[a%
%
%
]{Jason Bonnette}
\author[b%
%
%
]{Arvid R. Boe}
\author[c%
%
%
]{Phillip A. Fay}
\author[d%
%
%
]{Felix B. Fritschi}
\author[e%
%
%
]{David B. Lowry}
\author[f%
%
%
]{Robert B. Mitchell}
\author[g%
%
%
]{Francis M. Rouquette Jr}
\author[h%
%
%
]{Yanqi Wu}
\author[a%
%
%
]{Arbel Harpak}
\author[a%
%
,\correspond%
]{Thomas E. Juenger}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects (G), environmental
(e.g., weather) cues, and their interaction. Measuring GxWeather
interactions can aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of cloned individuals from these populations at eight
sites across the central United States. Some genetic effects covary with
interpretable weather-based cues. For example, in the Gulf population,
65\% of genetic effects on the timing of vegetative growth covary with
daylength 14 days prior to green-up date, and 33\% of genetic effects on
the timing of flowering covary with cumulative rainfall in the seven
days prior to flowering. However, most variation in genetic effects is
site-specific and cannot be attributed to variation in weather
variables. We demonstrate that we can identify genetic variation with
GxWeather and assign these loci to specific weather-based cues or other
patterns. We predict that selective breeding for particular alleles at
the GxWeather loci could alter flowering responsiveness in a photoperiod
or rainfall-specific way. More broadly, our approach proposes a refined
characterization of genotype-by-environment interactions in any species
phenotyped in multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[17]{2}

@@

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (6--13), a key output of
selection driving adaptation to local environments (2, 14--17), and a
selection target for crop improvement to adapt crops to local or future
environments (18). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (14, 19--23) and other crops (24, 25). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (11). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (26). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes (27--30), previous empirical work has found limited evidence of
antagonistic pleiotropy (15, 31, 32). However, this work has been
limited by a known statistical bias that reduced detection of genetic
effects that differ in sign (31, 33, 34). Thus, despite substantial
interest in the frequencies of various forms of GxE, the frequency of
sign-changing GxE relative to other forms of GxE remains unknown.

@@

We expected patterns of GxE and GxWeather to vary at the locus,
subpopulation, and trait level. Thus, we were interested in modeling the
different ways effects might covary across our eight common gardens. We
jointly re-estimated the genetic effects of a subset of SNPs across all
eight common gardens using \emph{mash} models, which can flexibly
capture many modes of covariance across gardens (SI Appendix, Sections
S1-S5, Datasets 1-6). To obtain genetic effects, we used subsets of
effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3;
Table S1). To obtain covariance structures, we specified both covariance
structures introduced in the initial \emph{mash} manuscript, and
GxWeather covariance matrices estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1; SI Appendix, Section S1). Then, we used a
greedy algorithm to select covariance matrices from each model's set
that significantly improved the model likelihood (SI Appendix, Section
S4; Table S2). Finally, we fit \emph{mash} models using the selected set
of covariance matrices on a subset relatively unlinked
(r\textsuperscript{2} \textless{} 0.2), `strong' genetic effects with
low \emph{p}-values (SI Appendix, Section S5).

@@

For green-up date, SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the \emph{mash} model of Gulf green-up fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of green-up. The covariance matrix for this weather
cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A).
\emph{Mash} models of the timing of vegetative growth in the Midwest
subpopulation and Both subpopulations did not include this GxWeather
covariance. The Midwest had 0.15\% weight on a GxWeather covariance
matrix of cumulative GDD for the 28d prior to green-up, a GxWeather
matrix that was visually similar to the Midwest's observed pattern of
phenotypic correlation (Figure~\ref{fig-map} B). Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations observed in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
subpopulations had substantially more mass on GxWeather covariance
matrices than the Midwest population (Figure~\ref{fig-covar} C) for the
onset of vegetative growth.

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither of these
covariance matrices significantly improved the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six \emph{mash} models of
strong effects, the GxWeather covariance matrices captured a minority of
the posterior weights of the strong effects (Figure~\ref{fig-covar}
C,E); in these five models, the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and single effect matrices with garden-specific effects.

@@

We expected to observe common rank-changing GxE at the level of
individual loci as this is a key theoretical prediction of local
adaptation (27--30). To determine the frequency of rank-changing GxE, we
used the local false sign rate (lfsr), an analogue of the local false
discovery rate that establishes confidence in the effect sign, not the
effect's difference from zero, to determine significance. We required
lfsr significance (p \textless{} 0.05) in both gardens to include
effects. This means that our tests for a sign change between gardens
carry an equal statistical burden to those for effects with the same
sign. We separated kinds of effects at the level of individual loci into
SNP effects that differ in sign between gardens (effects with
rank-changing GxE between gardens), SNP effects that differ in magnitude
(effects that are large in one garden or region and smaller in others),
and SNP effects that are indistinguishable in two gardens (similarly
large or small in comparisons between gardens) (Figure~\ref{fig-effects}
A).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
with one or more homologs from rice or \emph{A. thaliana} with
functionally validated roles in flowering (SI Appendix, Dataset 7); the
QTL on Chr04K overlapped a gene recently functionally validated for
flowering in switchgrass (43). All flowering and green-up QTL intervals
contained at least one SNP significant in at least one \emph{mash} run
at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail
of significance, whichever was stricter (SI Appendix, Dataset 8). We
also looked for enrichments of \emph{mash} SNPs in the 1\% tail of
significance (the `\emph{mash} 1\% tail') within each QTL interval. At
the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
tail. Overall, there were five significant enrichments (p \textless{}
0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
QTL intervals. Thus, we were able to experimentally support the genomic
windows of some re-estimates of significant SNP effects with a QTL
mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We assigned genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to unmeasured, site-based patterns. We used this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative and reproductive growth
revealed many alleles with rank-changing GxE, or sign changes in their
effects between the Texas and North gardens (Figure~\ref{fig-effects}
B). The Gulf \& Both subpopulations had rank-changing GxE for green-up
date, while the Midwest subpopulation had rank-changing GxE for
flowering date. As phenological timings are major components of plant
fitness, this result supports theoretical models that local adaptation
should involve trade-offs at the level of individual loci (27--30). In
addition, it advances local adaptation research by using more than two
field sites and a wide range of genetic variation to determine that
rank-changing GxE for phenological traits is common in small genomic
regions (15, 44, 45).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related genetic variation, in that
flowering timing covaries with a day length change signal two days
before flowering occurs. In contrast, the Gulf subpopulation does not
have genetic variation in flowering that covaries with a photoperiod
cue. Instead, the Gulf subpopulation has genetic variation in flowering
that covaries with the rainfall that occurs in the week prior to
flowering. Three genomic regions affecting flowering that we
re-estimated across all eight sites were also supported by QTL from an
independent mapping population at these sites (Figure~\ref{fig-qtl} B).

@@

\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
specified in \emph{mash} and the posterior weights placed on all
covariance matrices. (a) Common gardens are arranged in latitudinal
order within the matrices. Top row: Four example canonical covariance
matrices. Canonical matrices (purple) have simple interpretations, such
as equal effects across all common gardens, or effects specific to a
single common garden. Bottom row: Five example GxWeather covariance
matrices specified for the green-up date or flowering date phenotype;
these matrices were created from environment-specific correlations
across eight common gardens, and are described in Table 1. (b,d) Total
posterior weight placed on each covariance matrix type specified for (b)
green-up date and (d) flowering date \emph{mash} models, within and
between two genetic subpopulations. Covariance matrices included in
\emph{mash} that had zero posterior weight in all three \emph{mash} runs
on the genetic subpopulations, such as the identity matrix, are not
shown. (c,e) Total posterior weight placed on covariance matrices that
were hypothesized or canonical, for the (c) green-up date phenotype and
(e) flowering date phenotype.}

@@

\includegraphics{images/Figure_3_Fraction_GxE_by_Region_six_midwest_sites.png}

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05 at each pair of sites contrasted. a) Examples of
effect patterns at three pairs of sites with three types of GxE. All
effects are for an alternate allele, with the reference allele effect
defined to be zero at both gardens and represented by the dashed
vertical line. Sign: Effects that differ in sign at these pairs of
gardens (p \textless{} 0.05, lfsr). Magnitude: Effects identical in sign
(p \textless{} 0.05, lfsr) that differ in magnitude by a factor of
\textgreater0.4. Not Distinguishable: Effects not distinguishable by
magnitude nor sign of the effect, with no measurable GxE. b) The
fraction of effects with each GxE type for the onset of vegetative
growth (green-up date) and reproductive growth (flowering date), within
and between two genetic subpopulations. Common gardens are grouped by
the larger region they came from: North gardens are within the natural
range of the Midwest subpopulation, and include MO, NE, MI, and SD,
while Texas gardens are within the natural range of the Gulf
subpopulation, and include TX1, TX2, and TX3.}

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (38, 40, 45). We scored the
onset of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The onset of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis.

The formation and resequencing of the diversity panel has been described
previously (38). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (46), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both' subpopulations). Analyses to determine
narrow-sense heritability (h\textsuperscript{2}) for green-up and
flowering were done using linear mixed models and followed (38) (SI
Appendix, Section S6).

@@

We were interested in specifying genetic models for trait variation that
allowed more than one form of GxE, as we reasoned that different loci
should display different forms of GxE. In addition, we were interested
in an unbiased estimation of the frequency of rank-changing GxE relative
to other forms of GxE, as the presence of rank-changing GxE at the level
of individual loci is a key theoretical prediction of local adaptation.
\emph{Mash} allowed us to both specify multiple forms of GxE and
GxWeather and conduct unbiased statistical tests for when SNP effects
changed sign between common gardens.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Philip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, rank-changing
genotype-by-environment interaction, photoperiod, cumulative
rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author[a%
,\equalcont%
,\correspond%
]{Alice H. MacQueen}
\author[a%
,\equalcont%
%
]{Li Zhang}
\author[a%
,\equalcont%
%
]{Samuel A. Smith}
\author[a%
%
%
]{Jason Bonnette}
\author[b%
%
%
]{Arvid R. Boe}
\author[c%
%
%
]{Philip A. Fay}
\author[d%
%
%
]{Felix B. Fritschi}
\author[e%
%
%
]{David B. Lowry}
\author[f%
%
%
]{Robert B. Mitchell}
\author[g%
%
%
]{Francis M. Rouquette Jr}
\author[h%
%
%
]{Yanqi Wu}
\author[a%
%
%
]{Arbel Harpak}
\author[a%
%
,\correspond%
]{Thomas E. Juenger}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depend on genetic effects (G), environmental
(e.g., weather) cues, and their interaction. Here, we measure
phenological timings in two highly divergent switchgrass (\emph{Panicum
virgatum}) subpopulations using repeated plantings of cloned individuals
at eight sites across the central United States. The timing of
vegetative growth for the two subpopulations reversed between their two
natural ranges and had strong negative correlations between these
regions; in contrast, the timing of flowering was positively~correlated
between gardens. We expect that these phenotypic correlations consist of
polygenic effects on phenology which have distinct patterns of GxE
segregating at different mapped loci. Thus, we infer the mixture of ways
genetic effects impact phenological timings, such as across common
gardens (GxE) or with weather cues (GxWeather). We demonstrate that we
can identify genetic variation with GxWeather and assign genetic loci to
specific weather-based cues or other patterns. For example, in the Gulf
subpopulation, 65\% of genetic effects on the timing of vegetative
growth covary with daylength 14 days prior to green-up date, and 33\% of
genetic effects on the timing of flowering covary with cumulative
rainfall in the week prior to flowering. However, most variation in
genetic effects cannot be attributed to variation in weather variables.
Selective breeding for particular alleles at GxWeather loci could alter
flowering responsiveness in a photoperiod or rainfall-specific way. More
broadly, our approach refines the characterization of
genotype-by-environment interactions and can be implemented in any
species phenotyped in multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[13]{2}

@@

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (6--13), a key output of
selection driving adaptation to local environments (2, 14--17), and a
selection target for crop improvement to adapt crops to local or future
environments (18). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (14, 19--23) and other crops (24, 25). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (11). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (26). Additionally, despite
theoretical predictions that local adaptation should involve trade-offs
caused by antagonistic pleiotropy, or alleles with effects with opposing
fitness outcomes (27--30), previous empirical work has found limited
evidence of trade-offs caused by this form of GxE (15, 31, 32). However,
this work has been limited by a known statistical bias that reduced
detection of genetic effects that differ in sign (31, 33, 34). Thus,
despite substantial interest in the frequencies of various forms of GxE,
the frequency of sign-changing GxE relative to other forms of GxE
remains unknown.

@@

We expected patterns of GxE and GxWeather to vary at the locus,
subpopulation, and trait level. Thus, we were interested in modeling the
different ways polygenic effects might covary across our eight common
gardens(43). We jointly re-estimated the genetic effects of a subset of
SNPs across all eight common gardens using \emph{mash} models, which can
flexibly capture many modes of covariance across gardens (SI Appendix,
Sections S1-S5, Datasets 1-6). To obtain genetic effects, we used
subsets of effects from garden- and subpopulation GWAS (SI Appendix,
Section S2-S3; Table S1). To obtain covariance structures, we specified
both covariance structures introduced in the initial \emph{mash}
manuscript (e.g., garden-specific effects; equal effects at all gardens,
Figure~\ref{fig-covar} A), and GxWeather covariance matrices estimated
from the covariance of empirical weather patterns at each garden at
specific times before the phenological event (Table 1; SI Appendix,
Section S1). Then, we used a greedy algorithm to select covariance
matrices from each model's set that significantly improved the model
likelihood (SI Appendix, Section S4; Table S2). Finally, we fit
\emph{mash} models using the selected set of covariance matrices on a
subset of relatively unlinked (r\textsuperscript{2} \textless{} 0.2),
`strong' genetic effects with low \emph{p}-values (SI Appendix, Section
S5).

@@

For green-up date, SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the \emph{mash} model of Gulf green-up fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of green-up. The covariance matrix for this weather
cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A).
\emph{Mash} models of the timing of vegetative growth in the Midwest
subpopulation and Both subpopulations did not include this GxWeather
covariance. The Midwest had 0.15\% weight on a GxWeather covariance
matrix of cumulative GDD for the 28d prior to green-up, a GxWeather
matrix that was visually similar to the Midwest's observed pattern of
phenotypic correlation (Figure~\ref{fig-map} B). Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in the
day prior to green-up (Figure~\ref{fig-covar} B). The average
temperature covariance matrix had negative covariances between Texas and
North gardens, though not as strong as the negative phenotypic
correlations observed in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
subpopulations had substantially more mass on GxWeather covariance
matrices than the Midwest population (Figure~\ref{fig-covar} C) for the
onset of vegetative growth.

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither of these
covariance matrices significantly improved the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six \emph{mash} models of
strong effects, the GxWeather covariance matrices captured a minority of
the posterior weights of the strong effects (Figure~\ref{fig-covar}
C,E); in these five models, the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and single effect matrices with garden-specific effects.
Re-estimation of genetic effects with \emph{mash} allowed us to identify
and quantify the fraction of loci exhibiting GxWeather patterns; a
minority of GxE was GxWeather covariance, and the majority of GxE
covariance in these models is driven by other, unknown drivers of GxE.

@@

We expected to observe common rank-changing GxE at the level of
individual loci as this is a key theoretical prediction of local
adaptation (27--30). Previous empirical work has found limited evidence
of trade-offs caused by antagonistic pleiotropy; however, this work had
a known statistical bias reducing the detection of effects that differed
in sign (15, 31, 32). To determine the frequency of rank-changing GxE,
we used the local false sign rate (lfsr), an analogue of the local false
discovery rate that establishes confidence in the effect sign, not the
effect's difference from zero, to determine significance. We required
lfsr significance (p \textless{} 0.05) in both gardens to include
effects. This means that our tests for a sign change between gardens
carry an equal statistical burden to those for effects with the same
sign. We separated kinds of effects at the level of individual loci into
SNP effects that differ in sign between gardens (effects with
rank-changing GxE between gardens), SNP effects that differ in magnitude
(effects that are large in one garden or region and smaller in others),
and SNP effects that are indistinguishable in two gardens (similarly
large or small in comparisons between gardens) (Figure~\ref{fig-effects}
A).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
with one or more homologs from rice or \emph{A. thaliana} with
functionally validated roles in flowering (SI Appendix, Dataset 7); the
QTL on Chr04K overlapped a gene recently functionally validated for
flowering in switchgrass (44). All flowering and green-up QTL intervals
contained at least one SNP significant in at least one \emph{mash} run
at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail
of significance, whichever was stricter (SI Appendix, Dataset 8). We
also looked for enrichments of \emph{mash} SNPs in the 1\% tail of
significance (the `\emph{mash} 1\% tail') within each QTL interval. At
the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
tail. Overall, there were five significant enrichments (p \textless{}
0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
QTL intervals. Thus, we were able to experimentally support the genomic
windows of some re-estimates of significant SNP effects with a QTL
mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered through selection at individual loci. Here, we
demonstrate that we can associate multiple patterns of GxWeather with
specific genomic regions using a switchgrass diversity panel grown at
eight common gardens. We assigned genetic effects to both GxWeather
patterns with interpretable weather-based cues, and to unmeasured,
site-based patterns. We used this approach to study GxWeather for the
timings of vegetative and reproductive development in the deeply
genetically diverged Gulf and Midwest subpopulations of switchgrass.

Our analysis of the timing of vegetative and reproductive growth
revealed many alleles with rank-changing GxE, or sign changes in their
effects between the Texas and North gardens (Figure~\ref{fig-effects}
B). The Gulf \& Both subpopulations had rank-changing GxE for green-up
date, while the Midwest subpopulation had rank-changing GxE for
flowering date. As phenological timings are major components of plant
fitness, this result supports theoretical models that local adaptation
should involve trade-offs due to antagonistic pleiotropy at the level of
individual loci (27--30). Experimental designs in local adaptation
research now often use more than two field sites and a wide range of
genetic variation; \emph{mash} and the local false sign rate facilitate
the analysis of these experimental designs. We use \emph{mash} to
determine that rank-changing GxE for phenological traits is common in
small genomic regions (15, 45, 46).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related genetic variation, in that
flowering timing covaries with a day length change signal two days
before flowering occurs. In contrast, the Gulf subpopulation does not
have genetic variation in flowering that covaries with a photoperiod
cue. Instead, the Gulf subpopulation has genetic variation in flowering
that covaries with the rainfall that occurs in the week prior to
flowering. Three genomic regions affecting flowering that we
re-estimated across all eight sites were also supported by QTL from an
independent mapping population at these sites (Figure~\ref{fig-qtl} B).
Models combining both subpopulations showed less signal, perhaps due to
the distinctness of cues in both subpopulations or to confounding caused
by population structure.

@@

\caption{\label{fig-covar}Example Canonical and GxWeather covariance
matrices specified in \emph{mash} and the posterior weights placed on
all covariance matrices. (a) Common gardens are arranged in latitudinal
order within the matrices. Top row: Four example canonical covariance
matrices. Canonical matrices (purple) have simple interpretations, such
as equal effects across all common gardens, or effects specific to a
single common garden. Bottom row: Five example GxWeather covariance
matrices specified for the green-up date or flowering date phenotype;
these matrices were created from environment-specific correlations
across eight common gardens, and are described in Table 1. (b,d) Total
posterior weight placed on each covariance matrix type specified for (b)
green-up date and (d) flowering date \emph{mash} models, within and
between two genetic subpopulations. Covariance matrices included in
\emph{mash} that had zero posterior weight in all three \emph{mash} runs
on the genetic subpopulations, such as the identity matrix, are not
shown. (c,e) Total posterior weight placed on covariance matrices that
were GxWeather or Canonical, for the (c) green-up date phenotype and (e)
flowering date phenotype.}

@@

\includegraphics{images/Figure_3_AH_comments.png}

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated SNP effects in eight common gardens, for effects
with lfsr \textless{} 0.05 at both gardens for each pair of gardens
contrasted. a) Examples of effect patterns at three pairs of sites with
three types of GxE. All effects are for an alternate allele, with the
reference allele effect defined to be zero at both gardens and
represented by the dashed vertical line. Sign: Effects that differ in
sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
not distinguishable by magnitude nor sign of the effect, with no
measurable GxE. Confidence intervals are illustrative that the effect
estimate does not overlap zero. b) The fraction of effects with each GxE
type for the onset of vegetative growth (green-up date) and reproductive
growth (flowering date), within and between two genetic subpopulations.
Common gardens are grouped by the larger region they came from: North
gardens are within the natural range of the Midwest subpopulation, and
include MO, NE, MI, and SD, while Texas gardens are within the natural
range of the Gulf subpopulation, and include TX1, TX2, and TX3.}

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (38, 40, 46). We scored the
onset of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The onset of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis.

The formation and resequencing of the diversity panel has been described
previously (38). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (47), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both' subpopulations). Analyses to determine
narrow-sense heritability (h\textsuperscript{2}) for green-up and
flowering were done using linear mixed models and followed (38) (SI
Appendix, Section S6).

@@

We were interested in specifying genetic models for trait variation that
allowed more than one form of GxE, as we reasoned that different loci
should display different forms of GxE (e.g., changes in effect magnitude
vs sign). In addition, we were interested in an unbiased estimation of
the frequency of rank-changing GxE relative to other forms of GxE, as
the presence of rank-changing GxE at the level of individual loci is a
key theoretical prediction of local adaptation. \emph{Mash} allowed us
to both specify multiple forms of GxE and GxWeather and conduct unbiased
statistical tests for when SNP effects changed sign between common
gardens.

To use \emph{mash} on our diversity panel, we had to specify both a
relatively uncorrelated set of covariance matrices, which in our case
defined types of GxE and GxWeather between gardens, and we had to
specify subsets of SNP effect estimates and standard errors for our
traits at each common garden. To specify a set of covariance matrices,
we first defined many covariance matrices, including GxWeather matrices
that represented the correlation in weather cues between gardens before
the phenological event (SI Appendix, Section S1), then implemented a
model selection approach that used a greedy algorithm to evaluate if the
log likelihood of the \emph{mash} model was significantly improved as
additional covariance matrices were included (SI Appendix, Section S4).
To specify subsets of SNP effect estimates and standard errors to use
for both the greedy algorithm and with the optimal set of covariance
matrices, we first calculated best linear unbiased predictors (BLUPs)
for each phenological trait in each genetic subpopulation and each
common garden (SI Appendix, Section S2). Next, we determined effect
estimates for 8.8 to 12.3 million SNPs per subpopulation by conducting
garden-specific GWAS on these BLUPs using \(k\) vectors of singular
values to correct for population structure, where \(k\) was the smallest
integer value that made the genomic control coefficient closest to 1 (SI
Appendix, Section S2). Singular values were computed using singular
value decomposition of the matrix of all SNPs, with iterative SNP
pruning and removal of regions in long-range linkage disequilibrium
(48). Third, to make the \emph{mash} models computationally feasible, we
extracted two subsets of SNP effect estimates and standard errors from
our GWAS effect estimates: (i) effects from a subset of ``strong'' tests
corresponding to stronger effects on our traits; (ii) results from a
random subset of all tests to correspond to an unbiased representation
of all effects (SI Appendix, Section S3). We used the subset of random
effects in our greedy \emph{mash} algorithm (SI Appendix, Section S4)
and used the subset of strong effects in our \emph{mash} models with the
optimal set of covariance matrices (SI Appendix, Section S5).
